---
title: "Animated plots"
author: "Limor Raviv"
date: "12 April"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How to make beautiful animated plots using gganimate and tweenr

### This code is for creating plots that build up over time (i.e., the x-axis)

```{r load packages, echo=TRUE, warning=FALSE}
library(ggplot2)
library(gganimate)
library(animation)
library(tweenr)
```

### First, load your data.

You data should include at least 4 coloumns:

1. A variable of interest  (e.g., accuracy scores), which will be the y-axis 

2. A variable representing time (e.g., trial number), which will be the x-axis 

3. A varaible representing individual data points (e.g., participant number) 

4. A variable representing condition to which individuals are assigned (e.g., male vs. female) 

```{r load data, include=TRUE}

data <- read.csv("animated_plot_data.csv") 

head(data,30) # show the first 30 rows of the data frame
```

Then, use the tweenr package to make a smooth transition between each time point.

The "tween_elements functions will create two additional coloums to your data frame: ".frame" (corresponding to the origina trial number) and ".group" (corresponding to the original participant number).

```{r smooth, include=TRUE}

data$ease <- "linear" #add a coloumn to the data frame

data2<- tween_elements(data, 'Trial.Number', 'Participant', 'ease', nframes = 100) # create 100 frames between each trial number

head(data2,30) # show the first 30 rows of the data frame
```

Now, we'll create our wanted plot using ggplot.


```{r plot, warning=FALSE}
plot<- ggplot(data=data2, aes(x = Trial.Number, y = Accuracy, group=.group, color=Condition, frame=.frame, cumulative=TRUE))+
  geom_line(lwd=1, alpha=0.8)+  #change line trasparency with alpha = x
  scale_x_continuous("Trial Number", breaks=c(1:17))+
  scale_y_continuous("Accuracy")+
  theme_classic()+
  theme(text = element_text(size = 25), axis.text.x = element_text(size=20), legend.position=c(.8,.2))

```

And finally, let's animate the plot and save it!


```{r animate, include=TRUE, fig.show="animate""}

# change the speed of the animation by chanding interval = x

gganimate(plot, interval = 0.05)

# remove the title by frame by adding title_frame = FALSE
gganimate(plot, title_frame = FALSE,filename="animated_plot.gif",ani.height=500, ani.width=700, interval = 0.05) # save as gif


```

